---
title: "p8105_hw2_hyw2105"
output: github_document
date: "2026-10-01"
---


```{r}
library(tidyverse)
```


## Problem1
```{r}
subway_df =
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    na = c("", "NA", "."),
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )
  ) |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

subway_df
```

The cleaned dataset contains information on each entrance and exit of NYC
subway stations, including the subway line, station name, station latitude
and longitude, the routes served (`route1` to `route11`), whether entry is
allowed, vending availability, entrance type, and ADA compliance. I imported
the data with relative paths, specified `route8` to `route11` as character
at import so that all route variables share the same type, cleaned the
variable names, retained the relevant variables, and converted `entry` from
a YES/NO character variable to a logical variable. The resulting dataset has
`r nrow(subway_df)` rows and `r ncol(subway_df)` columns.

These data are not tidy: route information is spread across 11 columns, so
the route number is stored in the column names rather than as its own
variable.
