---
title: "p8105_hw2_hyw2105"
output: github_document
date: "2026-10-01"
---


```{r}
library(tidyverse)
```


## Problem1
```{r}
subway_df =
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    na = c("", "NA", "."),
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )
  ) |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

subway_df
```

The cleaned dataset contains information on each entrance and exit of NYC
subway stations, including the subway line, station name, station latitude
and longitude, the routes served (`route1` to `route11`), whether entry is
allowed, vending availability, entrance type, and ADA compliance. I imported
the data with relative paths, specified `route8` to `route11` as character
at import so that all route variables share the same type, cleaned the
variable names, retained the relevant variables, and converted `entry` from
a YES/NO character variable to a logical variable. The resulting dataset has
`r nrow(subway_df)` rows and `r ncol(subway_df)` columns.

These data are not tidy: route information is spread across 11 columns, so
the route number is stored in the column names rather than as its own
variable.


## Problem1-question1: How many distinct stations?
```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```
There are 465 distinct stations.

## Problem1-question2: How many stations are ADA compliant?
```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```
There are 84 ADA compliant.

## Problem1-question3: What proportion of entrances / exits without vending allow entrance?
```{r}
subway_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```
The mean is 0.3770492.


## Reformat route information
To make route number and route name distinct variables, I used
`pivot_longer()` to convert `route1` to `route11` into two variables:
`route_number` and `route_name`. Because most entrances serve only a few
routes, many of the route columns were empty; I removed rows with missing
`route_name`, since these do not correspond to any actual route.
```{r}
subway_tidy_df =
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
  drop_na(route_name)

subway_tidy_df
```
The resulting dataset has 4270 rows and 10 columns, where each row represents one route
served by one station entrance or exit.


### Stations serving the A train

```{r}
n_a_stations =
  subway_tidy_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

n_a_ada_stations =
  subway_tidy_df |>
  filter(
    route_name == "A",
    ada == TRUE
  ) |>
  distinct(line, station_name) |>
  nrow()

n_a_stations
n_a_ada_stations
```

There are 60 distinct stations that serve the A train. Of
these, 17 are ADA compliant.


# Problem 2
## Import and clean Trash Wheel data
```{r}
mr_trash_df =
  readxl::read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N710"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    year = as.numeric(year),
    trash_wheel = "mr"
  )

mr_trash_df
```

##Professor Trash Wheel and Gwynns Falls
```{r}
professor_trash_df =
  readxl::read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M135"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    trash_wheel = "professor"
  )

professor_trash_df

gwynnda_trash_df =
  readxl::read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynns Falls Trash Wheel",
    range = "A2:L352"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    trash_wheel = "gwynnda"
  )

gwynnda_trash_df
```

There are 132 rows in professor_trash_df and 349 rows in gwynnda_trash_df.

## Combine Trash Wheel datasets
```{r}
trash_wheel_df =
  bind_rows(
    mr_trash_df,
    professor_trash_df,
    gwynnda_trash_df
  ) |>
  relocate(trash_wheel)

trash_wheel_df
```
There are 1,188 rows（707 + 132 + 349）and 15 columns，count(). And we can see there that mr has 707、professor has 132、gwynnda has 349.

##filter → pull → vector summary
```{r}
professor_total_weight =
  trash_wheel_df |>
  filter(trash_wheel == "professor") |>
  pull(weight_tons) |>
  sum(na.rm = TRUE)

gwynnda_june_2022_butts =
  trash_wheel_df |>
  filter(
    trash_wheel == "gwynnda",
    year == 2022,
    month == "June"
  ) |>
  pull(cigarette_butts) |>
  sum(na.rm = TRUE)

professor_total_weight
gwynnda_june_2022_butts
```

The data come from the September 2025 version of the Trash Wheel
Collection Data. For each trash wheel, I specified the sheet and the cell
range in `read_excel()` to skip the empty first row and exclude non-data
columns, cleaned the variable names, and removed rows without a dumpster
number (the grand total row at the bottom of each sheet). For Mr. Trash
Wheel, I rounded `sports_balls` to the nearest integer and converted `year`
from character to numeric so that it matches the other two sheets. I added
a `trash_wheel` variable to each dataset before combining them with
`bind_rows()`.

The combined dataset contains `r nrow(trash_wheel_df)` observations, where
each row is one dumpster collected by one of three trash wheels. Key
variables include the collection date (`month`, `year`, `date`), the
weight (`weight_tons`) and volume (`volume_cubic_yards`) of trash, and
counts of specific items such as `plastic_bottles` and `cigarette_butts`.
Some items were not recorded by every trash wheel: `sports_balls` is only
available for Mr. Trash Wheel, and `glass_bottles` is not recorded for
Gwynnda, so these values are missing for the corresponding rows. The
`month` variable also has some inconsistent formatting in the original
file (e.g., trailing spaces and lowercase entries).

For available data, Professor Trash Wheel collected a total of
282.26 tons of trash. Gwynnda collected a total of
18120 cigarette butts in June 2022.