---
title: "p8105_hw2_hyw2105"
output: github_document
date: "2026-10-01"
---


```{r}
library(tidyverse)
```


## Problem1
```{r}
subway_df =
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    na = c("", "NA", "."),
    col_types = cols(
      Route8 = col_character(),
      Route9 = col_character(),
      Route10 = col_character(),
      Route11 = col_character()
    )
  ) |>
  janitor::clean_names() |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    route1:route11,
    entry,
    vending,
    entrance_type,
    ada
  ) |>
  mutate(
    entry = case_match(
      entry,
      "YES" ~ TRUE,
      "NO" ~ FALSE
    )
  )

subway_df
```

The cleaned dataset contains information on each entrance and exit of NYC
subway stations, including the subway line, station name, station latitude
and longitude, the routes served (`route1` to `route11`), whether entry is
allowed, vending availability, entrance type, and ADA compliance. I imported
the data with relative paths, specified `route8` to `route11` as character
at import so that all route variables share the same type, cleaned the
variable names, retained the relevant variables, and converted `entry` from
a YES/NO character variable to a logical variable. The resulting dataset has
`r nrow(subway_df)` rows and `r ncol(subway_df)` columns.

These data are not tidy: route information is spread across 11 columns, so
the route number is stored in the column names rather than as its own
variable.


## Problem1-question1: How many distinct stations?
```{r}
subway_df |>
  distinct(line, station_name) |>
  nrow()
```
There are 465 distinct stations.

## Problem1-question2: How many stations are ADA compliant?
```{r}
subway_df |>
  filter(ada == TRUE) |>
  distinct(line, station_name) |>
  nrow()
```
There are 84 ADA compliant.

## Problem1-question3: What proportion of entrances / exits without vending allow entrance?
```{r}
subway_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```
The mean is 0.3770492.


## Reformat route information
To make route number and route name distinct variables, I used
`pivot_longer()` to convert `route1` to `route11` into two variables:
`route_number` and `route_name`. Because most entrances serve only a few
routes, many of the route columns were empty; I removed rows with missing
`route_name`, since these do not correspond to any actual route.
```{r}
subway_tidy_df =
  subway_df |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
  drop_na(route_name)

subway_tidy_df
```
The resulting dataset has 4270 rows and 10 columns, where each row represents one route
served by one station entrance or exit.


### Stations serving the A train

```{r}
n_a_stations =
  subway_tidy_df |>
  filter(route_name == "A") |>
  distinct(line, station_name) |>
  nrow()

n_a_ada_stations =
  subway_tidy_df |>
  filter(
    route_name == "A",
    ada == TRUE
  ) |>
  distinct(line, station_name) |>
  nrow()

n_a_stations
n_a_ada_stations
```

There are 60 distinct stations that serve the A train. Of
these, 17 are ADA compliant.


# Problem 2
## Import and clean Trash Wheel data
```{r}
mr_trash_df =
  readxl::read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N710"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    sports_balls = as.integer(round(sports_balls)),
    year = as.numeric(year),
    trash_wheel = "mr"
  )

mr_trash_df
```

##Professor Trash Wheel and Gwynns Falls
```{r}
professor_trash_df =
  readxl::read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M135"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    trash_wheel = "professor"
  )

professor_trash_df

gwynnda_trash_df =
  readxl::read_excel(
    "data/202509 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynns Falls Trash Wheel",
    range = "A2:L352"
  ) |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(
    trash_wheel = "gwynnda"
  )

gwynnda_trash_df
```

There are 132 rows in professor_trash_df and 349 rows in gwynnda_trash_df.

## Combine Trash Wheel datasets
```{r}
trash_wheel_df =
  bind_rows(
    mr_trash_df,
    professor_trash_df,
    gwynnda_trash_df
  ) |>
  relocate(trash_wheel)

trash_wheel_df
```
There are 1,188 rows（707 + 132 + 349）and 15 columns，count(). And we can see there that mr has 707、professor has 132、gwynnda has 349.

##filter → pull → vector summary
```{r}
professor_total_weight =
  trash_wheel_df |>
  filter(trash_wheel == "professor") |>
  pull(weight_tons) |>
  sum(na.rm = TRUE)

gwynnda_june_2022_butts =
  trash_wheel_df |>
  filter(
    trash_wheel == "gwynnda",
    year == 2022,
    month == "June"
  ) |>
  pull(cigarette_butts) |>
  sum(na.rm = TRUE)

professor_total_weight
gwynnda_june_2022_butts
```

The data come from the September 2025 version of the Trash Wheel
Collection Data. For each trash wheel, I specified the sheet and the cell
range in `read_excel()` to skip the empty first row and exclude non-data
columns, cleaned the variable names, and removed rows without a dumpster
number (the grand total row at the bottom of each sheet). For Mr. Trash
Wheel, I rounded `sports_balls` to the nearest integer and converted `year`
from character to numeric so that it matches the other two sheets. I added
a `trash_wheel` variable to each dataset before combining them with
`bind_rows()`.

The combined dataset contains `r nrow(trash_wheel_df)` observations, where
each row is one dumpster collected by one of three trash wheels. Key
variables include the collection date (`month`, `year`, `date`), the
weight (`weight_tons`) and volume (`volume_cubic_yards`) of trash, and
counts of specific items such as `plastic_bottles` and `cigarette_butts`.
Some items were not recorded by every trash wheel: `sports_balls` is only
available for Mr. Trash Wheel, and `glass_bottles` is not recorded for
Gwynnda, so these values are missing for the corresponding rows. The
`month` variable also has some inconsistent formatting in the original
file (e.g., trailing spaces and lowercase entries).

For available data, Professor Trash Wheel collected a total of
282.26 tons of trash. Gwynnda collected a total of
18120 cigarette butts in June 2022.


# Problem 3

## Baseline demographics
```{r}
mci_baseline_df =
  read_csv(
    "data/data_mci/MCI_baseline.csv",
    skip = 1,
    na = c("", "NA", ".")
  ) |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      0 ~ "female"
    ),
    apoe4 = case_match(
      apoe4,
      1 ~ "carrier",
      0 ~ "non-carrier"
    )
  )

mci_baseline_df
```

```{r}
mci_baseline_eligible_df =
  mci_baseline_df |>
  filter(
    is.na(age_at_onset) | age_at_onset > current_age
  )

mci_baseline_eligible_df
```

##How many participants were recruited, and of these how many develop MCI? 
##What is the average baseline age? What proportion of women in the study are APOE4 carriers?
```{r}
n_recruited =
  nrow(mci_baseline_df)

n_eligible =
  nrow(mci_baseline_eligible_df)

n_develop_mci =
  mci_baseline_eligible_df |>
  filter(!is.na(age_at_onset)) |>
  nrow()

mean_baseline_age =
  mci_baseline_eligible_df |>
  pull(current_age) |>
  mean()

n_women =
  mci_baseline_eligible_df |>
  filter(sex == "female") |>
  nrow()

n_women_carrier =
  mci_baseline_eligible_df |>
  filter(
    sex == "female",
    apoe4 == "carrier"
  ) |>
  nrow()

prop_women_carrier =
  n_women_carrier / n_women

n_recruited
n_eligible
n_develop_mci
mean_baseline_age
prop_women_carrier
```

The baseline file contains a first line describing the coding of each
variable, so I skipped it at import. Missing values for age at MCI onset
are coded as `.` in the raw file, so I included `.` as a missing value
code. I cleaned the variable names and recoded `sex` (1 = male, 0 =
female) and `apoe4` (1 = carrier, 0 = non-carrier) from numeric codes to
descriptive character values, following the coding description in the
file. To meet the inclusion criterion of no MCI at baseline, I removed
participants whose age at MCI onset was less than or equal to their
baseline age.

A total of 483 participants were recruited. After removing 4 participants who did not meet the inclusion criterion, 479 participants remained, and 93 of them developed MCI during follow-up. The average baseline age was 65 years, and 30% of women in the study were APOE4 carriers.

## Amyloid biomarker data
```{r}
mci_amyloid_df =
  read_csv(
    "data/data_mci/mci_amyloid.csv",
    skip = 1,
    na = c("", "NA", ".", "Na")
  ) |>
  janitor::clean_names() |>
  rename(id = study_id) |>
  pivot_longer(
    baseline:time_8,
    names_to = "time",
    names_prefix = "time_",
    values_to = "amyloid_ratio"
  ) |>
  mutate(
    time = replace(time, time == "baseline", "0"),
    time = as.numeric(time)
  )

mci_amyloid_df
```

Like the baseline file, the amyloid file begins with a line describing the
variables, so I skipped it at import. Missing values in this file are coded
inconsistently as both `NA` and `Na`; because `Na` was not initially
recognized as missing, all amyloid columns were read as character. I
added `Na` to the missing value codes at import, which resolved this issue.
I cleaned the variable names, renamed `study_id` to `id` to match the
baseline data, and used `pivot_longer()` to convert the five measurement
columns (baseline and years 2, 4, 6, and 8) into a `time` variable and an
`amyloid_ratio` variable. The baseline visit was recoded as time 0, and
`time` was converted to numeric.

The tidy amyloid dataset has 2435 rows and 3 columns, with each row representing one
participant at one visit. Some amyloid measurements are missing, and these were retained.


## Compare participants across datasets
```{r}
baseline_only_df =
  anti_join(
    mci_baseline_eligible_df,
    mci_amyloid_df,
    by = "id"
  )

baseline_only_df

amyloid_only_df =
  anti_join(
    mci_amyloid_df,
    mci_baseline_eligible_df,
    by = "id"
  ) |>
  distinct(id)

amyloid_only_df

amyloid_not_recruited_df =
  anti_join(
    amyloid_only_df,
    mci_baseline_df,
    by = "id"
  )

amyloid_not_recruited_df
```
A total of 8 eligible participants appear in the baseline data but have no amyloid measurements, so no biomarker trajectory is available for them. Conversely, 16 participants appear in the amyloid data but not in the eligible baseline data. Of these, 4 were recruited but excluded because their MCI onset occurred at or before the study baseline, and 12 (IDs 484 to 495) do not appear in the baseline data at all, so their demographic information is unavailable. Because demographic and biomarker information are both needed to study biomarker trajectories, only participants who appear in both datasets are retained in the combined data below.

